---
title: "Data"
---

```{r setup, include = FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(dbplyr))
library(readxl)
```

# Introduction of the datasets

## Geo
The data for analysis is obtained by merging several Alzheimer's Disease related datasets.

The first dataset `death_rate` contains the mortality of Alzheimer’s Disease in each state from 2014 to 2021, which can be accessed [here](https://www.cdc.gov/nchs/pressroom/sosmap/alzheimers_mortality/alzheimers_disease.htm).
```{r, message = FALSE, warning = FALSE}
death_rate = 
  read_csv("./Data/death_rate.csv") |>
  janitor::clean_names() |>
  rename(death_rate = rate) |>
  select(state, year, death_rate) |>
  filter(year >= 2018) |>
  arrange(state, year)
```

`econdata` dataset contains the economic status of each state from 2008 to 2021, including GDP, income, price index, expenditure on healthcare etc. The data is downloaded from [Bureau of Economic Analysis](https://apps.bea.gov/iTable/?reqid=70&step=1&acrdn=2&_gl=1*14uawee*_ga*MzQ4ODY3MjA1LjE3MDEzMTIzMDc.*_ga_J4698JNNFT*MTcwMTMxMjMwNi4xLjEuMTcwMTMxMjM4OS4wLjAuMA..#eyJhcHBpZCI6NzAsInN0ZXBzIjpbMSwyOSwyNSwzMSwyNiwyNywzMF0sImRhdGEiOltbIlRhYmxlSWQiLCI2MDAiXSxbIk1ham9yX0FyZWEiLCIwIl0sWyJTdGF0ZSIsWyIwIl1dLFsiQXJlYSIsWyJYWCJdXSxbIlN0YXRpc3RpYyIsWyItMSJdXSxbIlVuaXRfb2ZfbWVhc3VyZSIsIkxldmVscyJdLFsiWWVhciIsWyItMSJdXSxbIlllYXJCZWdpbiIsIi0xIl0sWyJZZWFyX0VuZCIsIi0xIl1dfQ==).
```{r, message = FALSE, warning = FALSE}
econ_data = 
  read_csv("./Data/econdata.csv") |>
  janitor::clean_names() |>
  select(geo_name, description, x2018:x2021) |>
  mutate(description = 
           case_match(
             description,
             "GDP"~"gdp",
             "PCE"~"pce",
             "Regional price parities"~"rpp",
             "employment"~"employment",
             "healthcare expenditure"~"hc_exp",
             "personal income"~"income",
             "population"~"population")) |>
  pivot_longer(
    x2018:x2021,
    names_to = "year",
    values_to = "value"
  ) |>
  pivot_wider(
    names_from = "description",
    values_from = "value"
  ) |>
  rename(state = geo_name) |>
  mutate(year = as.numeric(str_replace(year, "x", "")),
         gdp = gdp/(population/1000000),
         pce = pce/(population/1000000),
         employment = round(employment/(population), 4),
         hc_exp = hc_exp/(population/1000000),
         income = income/(population/1000000),
         ) |>
  select(-population) |>
  arrange(state, year)
```

`edu_level` dataset contains the percentage of population obtained Bachelor's Degree or Higher in each state from 2018 to 2021. The data is extracted from [Federal Reserve Economic Data](https://fred.stlouisfed.org/release/tables?rid=330&eid=391444&od=2018-01-01#).
```{r, message = FALSE, warning = FALSE}
edu_level = 
  read_excel("./Data/edu_level.xlsx") |>
  janitor::clean_names() |>
  select(state, x2018:x2021) |>
  pivot_longer(
    x2018:x2021,
    names_to = "year",
    values_to = "edu_level"
  ) |>
  mutate(year = as.numeric(str_replace(year, "x", "")),
         edu_level = edu_level/100) |>
  arrange(state, year)
```

`diabetes` dataset contains the rate of diabetes by state from 2018 to 2021. The data can be accessed [here](https://www.kff.org/other/state-indicator/adults-with-diabetes/?currentTimeframe=1&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D).
```{r, message = FALSE, warning = FALSE}
diabetes_rate = 
  read_excel("./Data/diabetes.xlsx", range = "A1:C205") |>
  janitor::clean_names() |>
  rename(diabetes_rate = diabetes_proportion) |>
  mutate(diabetes_rate = round(as.numeric(diabetes_rate), 3),
         year = as.numeric(year)) |>
  arrange(state, year)
```

`Cardiovascular Disease` dataset contains the Cardiovascular Disease rate, and the data can be accessed [here](https://www.kff.org/other/state-indicator/percent-of-adults-with-cardiovascular-disease-by-sex/?currentTimeframe=1&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D).
```{r, message = FALSE, warning = FALSE}
cardio_rate = 
  read_excel("./Data/Cardiovascular Disease.xlsx", range = "A1:E52") |>
  janitor::clean_names() |>
  pivot_longer(
    x2018:x2021,
    names_to = "year",
    values_to = "cardio_rate"
  ) |>
  mutate(year = as.numeric(str_replace(year, "x", ""))) |>
  arrange(state, year)
```

`smoke` dataset contains the rate of smoking by state. The sources of the data are [CDC](https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm) and [KFF](https://www.kff.org/other/state-indicator/smoking-adults-by-sex/?currentTimeframe=2&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D).
```{r, message = FALSE, warning = FALSE}
smoke_rate =
  read_excel("./Data/smoke.xlsx") |>
  janitor::clean_names() |>
  pivot_longer(
    x2018:x2021,
    names_to = "year",
    values_to = "smoke_rate"
  ) |>
  mutate(year = as.numeric(str_replace(year, "x", "")),
         smoke_rate = smoke_rate/100) |>
  arrange(state, year)
```

`age_over_65` dataset contains the percentage of elderly people (>65 years old). This data is extracted from [Census Bureau](https://data.census.gov/table/ACSDP1Y2021.DP05?g=010XX00US$0400000&moe=false) and [PRB](https://www.prb.org/resources/which-us-states-are-the-oldest/).
```{r, message = FALSE, warning = FALSE}
elderly_rate = read_excel("./Data/age_over_65/age_over_65.xlsx") |>
  janitor::clean_names() |>
  pivot_longer(
    x2018:x2021,
    names_to = "year",
    values_to = "elder_rate"
  ) |>
  mutate(year = as.numeric(str_replace(year, "x", "")))|>
  arrange(state, year)
```

```{r}
obesity_rate = read_csv("./Data/obesity.csv") |>
  janitor::clean_names() |>
  pivot_longer(
    x2018:x2021,
    names_to = "year",
    values_to = "obesity_rate"
  ) |>
  mutate(year = as.numeric(str_replace(year, "x", "")))|>
  arrange(state, year)
```

These datasets were tidied and merged to generate
```{r, message = FALSE, warning = FALSE}
geo_data = left_join(death_rate, econ_data, by = c("state", "year")) |>
  left_join(edu_level, by = c("state", "year")) |>
  left_join(diabetes_rate, by = c("state", "year")) |>
  left_join(cardio_rate, by = c("state", "year")) |>
  left_join(smoke_rate, by = c("state", "year")) |>
  left_join(elderly_rate, by = c("state", "year")) |>
  left_join(obesity_rate, by = c("state", "year"))
```

```{r}
write.csv(geo_data, "./Data/geo_data.csv", row.names = FALSE)
```

## Clinical Trials data
Clinical Trials data is directly downloaded from **clinicaltrials.gov** for Alzheimer's Disease.
```{r read data, include = FALSE}
CT_data = 
  read_csv("./Data/ClinicalTrials_raw_data.csv") |>
  janitor::clean_names() 
```
There are a total of```r nrow(CT_data)```` and ```r length(CT_data)``` variables. Variables and description is listed below:

*```r colnames(CT_data)[1]```: 

*```r colnames(CT_data)[2]```: 

*```r colnames(CT_data)[3]```: 

*```r colnames(CT_data)[4]```:

*```r colnames(CT_data)[5]```:

*```r colnames(CT_data)[6]```:

*```r colnames(CT_data)[7]```:

*```r colnames(CT_data)[8]```:

*```r colnames(CT_data)[9]```:

*```r colnames(CT_data)[10]```:

*```r colnames(CT_data)[11]```:

# should simplify variables later 
