---
title: "Data"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
library(readxl)
library(tidyverse)
```

# Introduction of the datasets

## (AD) Demographic dataset
We extract data from ALZHEIMER’S ASSOCIATION REPORT to give an overview of current figures of Alzheimer's disease in the United States.

```{r upload dataset}
# Get demongraphic data
demo_data = 
  read_excel("./Data/Demographic.xlsx", range = "A1:D52") |>
  janitor::clean_names()
```
Demographic dataset contains the statistics about AD in ```r length(demo_data$state)``` states of the US. There are a total of ```r length(demo_data)``` variables. Variables and description is listed below:

*```r colnames(demo_data)[1]```: State name

*```r colnames(demo_data)[2]```: Projections of total numbers (in thousands) of Americans age 65 and older with Alzheimer's demential by state, 2020

*```r colnames(demo_data)[3]```: Number of deaths and annual mortality rate (per100,000) due to Alzheimer’s disease by state, 2019

*```r colnames(demo_data)[4]```: Number of caregivers of people with Alzheimer's or other dementias

## Clinical Trials data
Clinical Trials data is directly downloaded from **clinicaltrials.gov** for Alzheimer's Disease.
```{r read data, include = FALSE}
CT_data = 
  read_csv("./Data/ClinicalTrials_raw_data.csv") |>
  janitor::clean_names() 
```
There are a total of```r nrow(CT_data)```` and ```r length(CT_data)``` variables. Variables and description is listed below:

*```r colnames(CT_data)[1]```: 

*```r colnames(CT_data)[2]```: 

*```r colnames(CT_data)[3]```: 

*```r colnames(CT_data)[4]```:

*```r colnames(CT_data)[5]```:

*```r colnames(CT_data)[6]```:

*```r colnames(CT_data)[7]```:

*```r colnames(CT_data)[8]```:

*```r colnames(CT_data)[9]```:

*```r colnames(CT_data)[10]```:

*```r colnames(CT_data)[11]```:

# should simplify variables later 
