---
title: "Dena"
output: html_document
---
```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
library(ggplot2)
library(readxl)
```

The outcomes listed in Table 2 that we used in the artical include iADRS (Integrated Alzheimer's Disease Rating Scale), MMSE (Mini-Mental State Examination), and other clinical assessments. The table provides valuable insights into the efficacy of donanemab compared to placebo in slowing the progression of Alzheimer's disease over the 76-week period in the different populations studied.

The low/medium tau population included participants with mild cognitive impairment or early Alzheimer's disease who had low or intermediate levels of tau biomarkers in their cerebrospinal fluid (CSF). The high tau population included participants with early Alzheimer's disease who had high levels of tau biomarkers in their CSF.

The combined population included all participants in the trial, regardless of their tau biomarker levels. The purpose of dividing the participants into these populations was to assess the efficacy of donanemab in different subgroups of patients with Alzheimer's disease.

```{r}
iADRS <- read_excel("data/Donanemab_.xlsx", sheet = "1")

ggplot(iADRS, aes(x = as.factor(time), y = mean, fill = group, color = group)) +
  geom_point(position = position_dodge(0.9), size = 3, alpha = 0.7) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(0.9), width = 0.25) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Mean with 95% CI over Time",
       x = "Time",
       y = "Mean") +
  theme_minimal()
```

```{r}
CDR_SB <- read_excel("data/Donanemab_.xlsx", sheet = "2")

ggplot(CDR_SB, aes(x = as.factor(time), y = mean, fill = group, color = group)) +
  geom_point(position = position_dodge(0.9), size = 3, alpha = 0.7) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(0.9), width = 0.25) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Mean with 95% CI over Time",
       x = "Time",
       y = "Mean") +
  theme_minimal()
```

```{r}
ADCS_iADL <- read_excel("data/Donanemab_.xlsx", sheet = "3")

ggplot(ADCS_iADL, aes(x = as.factor(time), y = mean, fill = group, color = group)) +
  geom_point(position = position_dodge(0.9), size = 3, alpha = 0.7) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(0.9), width = 0.25) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Mean with 95% CI over Time",
       x = "Time",
       y = "Mean") +
  theme_minimal()
```

```{r}
ADCS_Cog13 <- read_excel("data/Donanemab_.xlsx", sheet = "4")

ggplot(ADCS_Cog13, aes(x = as.factor(time), y = mean, fill = group, color = group)) +
  geom_point(position = position_dodge(0.9), size = 3, alpha = 0.7) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(0.9), width = 0.25) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Mean with 95% CI over Time",
       x = "Time",
       y = "Mean") +
  theme_minimal()
```

```{r}
MMSE <- read_excel("data/Donanemab_.xlsx", sheet = "5")

ggplot(MMSE, aes(x = as.factor(time), y = mean, fill = group, color = group)) +
  geom_point(position = position_dodge(0.9), size = 3, alpha = 0.7) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(0.9), width = 0.25) +
  facet_wrap(~outcome, scales = "free_y") +
  labs(title = "Mean with 95% CI over Time",
       x = "Time",
       y = "Mean") +
  theme_minimal()
```

