---
title: "Preliminary Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
---

```{r, include = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Death rate by Regions

```{r}
data = read_csv("data/geo_data.csv") |> 
  mutate(region = case_when(
    state %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "New Jersey", "New York", "Pennsylvania", "Rhode Island", "Vermont") ~ "Northeast",
    state %in% c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin") ~ "Midwest",
    state %in% c("Alabama", "Arkansas", "Delaware", "District of Columbia", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia") ~ "South",
    state %in% c("Alaska", "Arizona", "California", "Colorado", "Hawaii", "Idaho", "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming") ~ "West"
  ))
```

In our endeavor to analyze Alzheimer's disease death rates across the United States, we have grouped the data by region, categorizing it into four distinct areas: Northeast, Midwest, South, and West. The average death rates from Alzheimer's disease vary across the different regions of the United States.

```{r}
data |>
  janitor::clean_names() |> 
  group_by(region) |> 
  summarise(avg_death_rate = mean(death_rate, na.rm = TRUE)) |> 
  arrange(avg_death_rate) |> 
  knitr::kable(digits = 3)
```

```{r}
data  |> 
  janitor::clean_names()  |> 
  group_by(region)  |> 
  ggplot(aes(x = fct_reorder(region, death_rate), y = death_rate, fill = region)) +
  geom_boxplot() +
  labs(title = "Boxplot of Death Rate by Regions", x = "Region", y = "Death Rate") +
  theme(axis.text.x = element_text(size = 8),
        plot.title = element_text(hjust = 0.5))
```

* The boxplot displays the distribution of Alzheimer's Disease death rates across different regions.

* Notably, the South region exhibits a wider spread of death rates, suggesting greater variability in Alzheimer's Disease mortality within this region compared to others.

* The observation of varying spreads among regions implies potential heterogeneity in Alzheimer's Disease death rates, indicating the importance of further investigation into the factors contributing to this variability.

## Death rate by (region, year) Combination

```{r}
data |>  
  group_by(region,year) |> 
  summarise(avg_by_year_region = mean(death_rate,na.rm = T)) |> 
  pivot_wider(
    names_from = region,
    values_from = avg_by_year_region
  ) |>  
  knitr::kable(digits = 3)
```

```{r}
data |>  
  group_by(region,year) |> 
  summarise(avg_by_year_region = mean(death_rate,na.rm = T)) %>% 
  ggplot(aes(x = year, y = avg_by_year_region, color = region)) +
  geom_line() + 
  geom_point() + 
  labs(
    y = "Average Alzheimer's Disease death rate",
    title = "Average Alzheimer's Disease death rate by Region, Year") +
  theme(plot.title = element_text(hjust = 0.5))
```

* The table and line graph above show the descriptive statistics and overall trend. It is clear that the Northeast region has a much lower and South region has a much higher average Alzheimer's Disease death rate than the other regions across the time period.

## Death Rate Comparison by Regions

From the boxplot above, we find that the Alzheimer’s Disease death rate in West region and Midwest region are almost overlapping. Though the median of Midwest region is higher, the variance in West region seems smaller. Thus, we want to study if the mean death rate in the two regions are significantly different.

Firstly, we check if two sets of data have equal variance

<center> *H~0~ : Equal  variance betweenWest and Midwest vs  H~1~ : Unequal  variance*</center>

```{r}
west_death_rate <- data |> 
  filter(region == "West") |> 
  pull(death_rate)

midwest_death_rate <- data |>  
  janitor::clean_names() |> 
  filter(region == "Midwest") |> 
  pull(death_rate)

var.test(west_death_rate,midwest_death_rate,alternative = "two.sided",conf.level = 0.95) |>  
  broom::tidy() |> 
  knitr::kable()
```

The result shows that we are confident to conclude that the variances are not equal. Next, we should perform 2 sample t.test with unknown and unequal variance.

```{r}
t.test(west_death_rate,midwest_death_rate,alternative = "less",conf.level = 0.95,paired = F,var.equal = FALSE ) |> 
  broom::tidy() |> 
  knitr::kable()
```

*   The results indicate that, based on the data, there is no significant difference in death rates between the "west" and "midwest" regions. The regional level may be too broad to discern significant differences. This finding suggests that variations in Alzheimer's mortality rates are potentially more nuanced and may become more apparent when examined at a smaller geographic scale. Therefore, for a more detailed and informative understanding of the distribution of Alzheimer's death rates, it would be beneficial to conduct further analysis at a more localized level, such as state or county.

# Death rate by Income Level

```{r}
data |> 
  group_by(year) |> 
  select(year,income) |> 
  summarise(median_income_by_year = median(income,na.rm = T)) |> 
  knitr::kable(digits = 3)
```

The table above shows the median personal annual income of each year. Then, we use the median personal annual income of each year to categorize different income groups (with"less than median income" as "Lower Income", "more than 1.3 times of median income" as "Upper Income", and those between these ranges as "Middle Income". And we use this income group division to calculate average Alzheimer's Disease death rate of each group.

```{r}
income_level_2018=
  data |>  
  group_by(year) |> 
  filter(year==2018) |> 
  mutate(
    income_group=case_when(
      income < 48986.21 ~"Lower Income",
      income > 1.3* 48986.21 ~"Upper Income",
      TRUE ~  "Middle income"
    )) 

income_level_2019=
  data |>  
  group_by(year) |> 
  filter(year==2019) |> 
  mutate(
    income_group=case_when(
      income <49792.03 ~"Lower Income",
      income > 1.3* 49792.03 ~"Upper Income",
      TRUE ~  "Middle income"
    ))

income_level_2020=
  data |>  
  group_by(year) |> 
  filter(year==2020) |> 
  mutate(
    income_group=case_when(
      income <52442.06 ~"Lower Income",
      income > 1.3* 52442.06 ~"Upper Income",
      TRUE ~  "Middle income"
    )) 
    
income_level_2021=
  data |>  
  group_by(year) |> 
  filter(year==2021) |> 
  mutate(
    income_group=case_when(
      income <54057.93 ~"Lower Income",
      income > 1.3* 54057.93 ~"Upper Income",
      TRUE ~  "Middle income"
    ))

income_level=
  bind_rows(income_level_2018,income_level_2019,income_level_2020,income_level_2021)

income_level |> 
  filter(!is.na(income_group)) |>  
  group_by(income_group) |> 
  summarise(avg_by_income = mean(death_rate,na.rm = T)) |>  
  arrange(avg_by_income) |> 
  knitr::kable(digits = 3)
```  

```{r}
income_level |> 
  filter(!is.na(income_group)) |>  
  group_by(income_group) |> 
  filter(!is.na(death_rate)) |> 
  ggplot(aes(x = fct_reorder(income_group,death_rate), y = death_rate,fill = income_group)) +
  geom_boxplot() +
  labs(
  title = "Boxplot of Alzheimer's Disease Death Rate by Income Groups", 
  x="Income Groups",
  y= "Alzheimer's Disease death rate") +
  theme(plot.title = element_text(hjust = 0.5))
```

* The boxes of different income groups are almost not overlapping with each other.

* The pattern is clear that people from lower income groups tend to have a higher Alzheimer's Disease Death Rate.

Then we perform statistical test to evluate the significance of difference of Alzheimer's Disease death rate among different income levels.

<center> *H~0~ : mean of death_rate are equal in three groups  vs  H~1~ :  at least two means are not equal* </center>

```{r}
anova_result <- aov(death_rate ~ factor(income_group), data = income_level)

TukeyHSD(anova_result) |>
  broom::tidy() |>
  knitr::kable()
```

* The ANOVA results indicate a significant difference in mean life expectancy among different income groups (p < 0.05). The post hoc tests further reveal:

  * A significant difference in mean life expectancy between Middle Income and Lower Income groups.

  * A significant difference in mean life expectancy between Upper Income and Lower Income groups.

  * No significant difference in mean life expectancy between Upper Income and Middle Income groups.

In summary, there are statistically significant variations in life expectancy across income groups, with specific differences observed in pairwise comparisons between Middle Income vs. Lower Income and Upper Income vs. Lower Income.

# Death rate by Employment

```{r}
data |> 
  group_by(year) |> 
  select(year,employment) |> 
  summarise(med_employ_by_year = median(employment,na.rm = T)) |> 
  knitr::kable(digits = 3)
```

The table above shows the median personal annual income of each year. Then, we use the median employment rate of each year to categorize different employment groups (with"less than or equal to median" as "Lower Employment Rate", "more than median" as "High Employment Rate". We use this employment group division to calculate average Alzheimer's Disease death rate of each group.

```{r}
employ_2018=
  data |>  
  group_by(year) |> 
  filter(year==2018) |> 
  mutate(
    employ_group=case_when(
      employment <= 0.617 ~"Low Employment Rate",
      employment >0.617~"High Employment Rate",
    )) 
  
employ_2019=
  data |>  
  group_by(year) |> 
  filter(year==2019) |> 
  mutate(
    employ_group=case_when(
      employment <= 0.617 ~"Low Employment Rate",
      employment > 0.617~"High Employment Rate",
    )) 

employ_2020=
  data |>  
  group_by(year) |> 
  filter(year==2020) |> 
  mutate(
    employ_group=case_when(
      employment <= 0.588 ~"Low Employment Rate",
      employment > 0.588~"High Employment Rate",
    )) 
    
employ_2021=
  data |>  
  group_by(year) |> 
  filter(year==2021) |> 
  mutate(
    employ_group=case_when(
      employment <= 0.612 ~"Low Employment Rate",
      employment >0.612~"High Employment Rate",
    )) 

employ_level=
  bind_rows(employ_2018,employ_2019,employ_2020,employ_2021)

employ_level |> 
  group_by(employ_group) |> 
  summarise(avg_by_employ = mean(death_rate,na.rm = T)) |>  
  knitr::kable(digits = 3)
```  
  
```{r}
employ_level |> 
  group_by(employ_group) |> 
  ggplot(aes(x = employ_group, y = death_rate,fill = employ_group)) +
  geom_boxplot() +
  labs(
  title = "Boxplot of Alzheimer's Disease Death Rate by Employment Rate",
  x="Employment Rate",
  y="Alzheimer's Disease Death Rate")
```

The calculation and pattern of box plot shows that low employment tends to a higher Alzheimer's Disease death rate. And we will use statistical test to check the significance of this difference.

<center> *H~0~ : mean high employment = mean low employment  vs  H~1~ : mean high employment ≠ mean low employment* </center>

```{r}
t.test(death_rate ~ factor(employ_group),data = employ_level, alternative = "less",conf.level = 0.95,paired = F,var.equal = TRUE ) |> 
  broom::tidy() |>
  knitr::kable()
```

* We have also performed 2 sample t.test to compare the mean death rate between the two groups. We conclude that the means are not equal. Since the method is similar to what we have used and displayed when studying the regional differences, we do not show the process here.

# Summary

In this analysis, we explored the impact of various factors on Alzheimer's Disease death rates using statistical tests and visualizations. Key findings include:

* Regional differences in Alzheimer's Disease death rates, with the South region having the highest and the Northeast region having the lowest average death rates.
* Significant variations in Alzheimer's Disease death rates based on income and employment groups, indicating that socioeconomic factors play a role in disease outcomes.